### 2019-03-26 4_implementation.tex
Apply Alex's comments to section 4.

One challenge in implementing Task Bench for such a wide variety of
systems is that the capabilities of the systems vary considerably. For
example, some systems are \emph{implicitly parallel}, and provide some
form of parallelism discovery from sequential programs, whereas others
are \emph{explicitly parallel} and require users to specify the
parallelism in the program. As a rule of thumb we aim for the Task
Bench implementations to be representative of how actual applications
would be written in the respective systems. Places where the Task
Bench implementations may be unidiomatic are noted below.

There are also aspects of some systems that we do not consider in our
evaluation. For example, a number of systems provide some level of
support for heterogeneous processors such as GPUs, but the support
varies so widely (if it exists at all) that it is challenging to come
up with a representative benchmark which allows these features to be
compared in an apples-to-apples manner.

In all cases, members of the respective programming systems' teams
were consulted during the development and evaluation of the
corresponding Task Bench implementations. Their insight was invaluable
in helping to ensure that we provide the highest level of quality in the Task
Bench implementations for each system.

@@

Chapel~\cite{Chapel07} is an explicitly parallel programming language
that follows a \emph{multi-resolution} approach to parallel
programming, with support for parallelism at a variety of
levels. Chapel's core features are a partitioned global address space
(PGAS) which permits remote access to objects from anywhere in the
machine, a data model for the distribution of data structures over the
global address space via \emph{domain maps}, tasks, and
synchronization primitives. Chapel thus permits a great deal of
flexibility in the choice of implementation strategy for applications.

For Task Bench, we decided to target a relatively low-level
implementation in Chapel, using explicit task instantiation (via
\lstinline[language=Chapel]{coforall}), bulk access to distributed
arrays for efficient data movement, and atomic integers for
synchronization. The advantage of this approach is that it permits
fine-tuning of the implementation for performance, and avoids relying
on advanced optimizations in the Chapel compiler, which might struggle
with the dynamic, input-dependent nature of the task graphs in Task
Bench.

Chapel also supports a more high-level approach to parallel
programming in which the granularities of tasks are determined
automatically by the compiler (via
\lstinline[language=Chapel]{forall}). This is the level that users
would typically target in their programs. The specific dependence
patterns supported by Task Bench might also be implementable at this
level, though a full Task Bench implementation would be challenging.

% FIXME: Does the Chapel compiler handle: bulk data movement, point-to-point synchronization

@@

Charm++~\cite{Charmpp93} is an explicitly parallel programming system
based on the actor model of distributed computation. Actors, or
\emph{chares}, are objects that live in their own address space and
can be migrated automatically between nodes for better load
balance. Chares are permitted to send messages to other chares via a
form of remote procedure call (RPC). Calling a method of another chare
is the primary form of synchronization and data movement in
Charm++. Additional support is provided for collective operations
between chares.

Our Task Bench implementation is straightforward: we create a chare
array for the task graph, with one chare for each column. Methods are
used to implement dependencies, and a task is executed as soon as its
dependencies are all available. Collectives are used to synchronize the
completion of the task graph execution.

@@

MPI~\cite{MPI} is the industry standard message-passing API for
programming supercomputers, and is the most widely used programming
model in high-performance scientific computing today. The goal of our
MPI implementation of Task Bench is to represent the common (if
unexciting) case which continues to dominate high-performance
computing: bulk synchronous programming with distinct computation and
communication phases. We use \lstinline[language=C++]{MPI_Isend} and
\lstinline[language=C++]{MPI_Irecv} to implement the communication
phase, posting receives before sends. Each task dependency maps to one
send/receive pair in MPI. Despite the relatively straightforward
design, our MPI implementation is among the best performing.

@@

PaRSEC is a task-based programming system which supports two distinct
programming models: \emph{parameterized task graphs}
(PTG)~\cite{PARSEC13} and \emph{dynamic task discovery}
(DTD)~\cite{PARSEC_DTD}.  PTG is a dataflow model in which programmers
write a concise, algebraic description of the tasks and dataflows in
the program. This compressed representation is expanded into a full
task graph by a source-to-source compiler. PTG is therefore a more
limited programming model, but can potentially provide better
performance because the graph description is amenable to static
analysis and can be pruned efficiently for better horizontal
scalability.

In the DTD model, tasks are enumerated in program order (by
dynamically executing the program), and dependencies between tasks are
identified automatically based on the input and output data of tasks. 
A task depends on another task if it reads data written by the other task,
and the data is copied automatically if the two tasks are executed on
different nodes. While the basic DTD model is implicitly parallel
because the system is responsible for the discovery of parallelism, in
order to achieve horizontal scalability, the program is executed in a
SPMD fashion on all nodes, and the user is responsible for eliding
tasks that are not directly connected to those that are to be executed
on the current node.

Data in PaRSEC is stored in a built-in matrix data structure with a 2D block cyclic distribution. 
The Task Bench implementation divides this matrix into $M$ by $N$ tiles where $N$ equals to the width of
the task graph and $M$ varies from 2 to the height of the task graph depending
on optimizations performed to alleviate write-after-read depdenencies. 
Tiles are distributed across nodes in pre-defined patterns, 
and can be passed to tasks on other nodes, but
can only be written on the node where the data is originally
allocated. In our Task Bench implementation, each task writes to a
tile as output, and these tiles are passed as inputs to the
subsequent sets of tasks. As in the OmpSs and OpenMP implementations,
the task launch code uses a switch statement as the PaRSEC API assumes
tasks have a fixed number of dependencies.

@@

Realm~\cite{Realm14} is an explicitly-parallel task-based programming
model that used internally by Legion~\cite{Legion12} and
Regent~\cite{Regent15}. Thus the Realm implementation of Task Bench can be
seen as a limit study of what can be achieved with Legion or Regent.

Tasks in Realm are executed in a deferred manner, with dependencies
between tasks explicitly defined by \emph{events} passed from one task
to another. Realm's data model supports \emph{regions} of memory which live on a single
node. Data must be explicitly copied, and the copies synchronized with
tasks via events.

@@

StarPU~\cite{StarPU11} is a task-based system that supports an DTD
programming model similar to PaRSEC. Our Task Bench implementation in
StarPU is very similar to PaRSEC, including the use of a switch
statement to handle dynamic numbers of task dependencies, and manual
elision of tasks on unrelated nodes to provide horizontal scalability.
A simplified version of the 2D block cyclic matrix from the Chameleon 
project~\cite{Chameleon} is used to describe the data produced by tasks in 
our Task Bench implementation.

>>>>

One challenge in implementing Task Bench for such a wide variety of
systems is that the capabilities of the systems vary considerably. For
example, some systems are \emph{implicitly parallel}, and provide some
form of parallelism discovery from sequential programs, whereas others
are \emph{explicitly parallel} and require users to specify the
parallelism in the program. The Task
Bench implementations are intended to reflect how actual applications
would be written in the respective systems. Places where the Task
Bench implementations may be unidiomatic are noted below.

There are also aspects of some systems that we do not consider in our
evaluation. For example, a number of systems provide some level of
support for heterogeneous processors such as GPUs, but the support
varies so widely (if it exists at all) that it is challenging to come
up with a representative benchmark that allows these features to be
compared in an apples-to-apples manner.

In all cases, members of the respective programming systems' teams
were consulted during the development and evaluation of the
corresponding Task Bench implementations. Where assistance was provided, the insights were invaluable
in helping to ensure that we provide the highest level of quality in the Task
Bench implementations for each system.

@@

Chapel~\cite{Chapel07} is an explicitly parallel programming language
that follows a \emph{multi-resolution} approach to parallel
programming, with support for parallelism at a variety of
levels. Chapel's core features are a partitioned global address space
(PGAS), a data model for the distribution of data structures, tasks, and
synchronization primitives. Certain operations can be lifted to entire arrays, enabling forms of data-parallelism. Chapel thus permits a great deal of
flexibility in the choice of implementation strategy for applications.

For Task Bench, we decided to target a low-level implementation in
Chapel, using explicit task instantiation (via
\lstinline[language=Chapel]{coforall}), bulk access to distributed
arrays for efficient data movement, and atomic integers for
synchronization. This approach is a good match for Task Bench, as the
Task Bench core provides implementations of the kernels.

@@

Charm++~\cite{Charmpp93} is an explicitly parallel programming system
based on the actor model. Actors, or
\emph{chares}, are objects that live in their own address space and
can be migrated automatically between nodes to
balance load. Chares send messages to other chares via a
form of remote procedure call (RPC). Calling a method of another chare
is the primary form of synchronization and data movement in
Charm++. Additional support is provided for collective operations
between chares.

Our Task Bench implementation is straightforward: we create a chare
array for the task graph, with one chare for each column of the task graph. Methods are
used to implement dependencies, and a task is executed as soon as its
dependencies are all available. Collectives are used to synchronize the
completion of the task graph execution.

@@

MPI~\cite{MPI} is the standard message-passing API for
programming supercomputers and remains the most widely used programming
model in high-performance scientific computing today. Our
MPI implementation of Task Bench represents the common case of bulk synchronous programs with distinct computation and
communication phases. We use \lstinline[language=C++]{MPI_Isend} and
\lstinline[language=C++]{MPI_Irecv} to implement the communication
phase, posting receives before sends. Each task dependency maps to one
send/receive pair in MPI. Despite the relatively straightforward
design, our MPI implementation is among the best performing.

@@

PaRSEC is a task-based programming system supporting two distinct
programming models: \emph{parameterized task graphs}
(PTG)~\cite{PARSEC13} and \emph{dynamic task discovery}
(DTD)~\cite{PARSEC_DTD}.  PTG is a dataflow model in which programmers
write a concise, algebraic description of the tasks and dataflows in
the program. This compressed representation is expanded into a full
task graph by a source-to-source compiler. PTG is therefore a more
limited programming model, but can potentially provide better
performance because the graph description is amenable to static
analysis and can be pruned efficiently for better horizontal
scalability.

In the DTD model, tasks are enumerated in program order (by
dynamically executing the program), and dependencies between tasks are
identified automatically based on the input and output data of tasks. 
A task depends on another task if it reads data written by the other task,
and the data is copied automatically if the two tasks are executed on
different nodes. While the basic DTD model is implicitly parallel
because the system is responsible for the discovery of parallelism, to
achieve horizontal scalability, the program is executed in a
SPMD fashion on all nodes, and the user is responsible for eliding
tasks that are not directly connected to those that are to be executed
on the current node.

Data in PaRSEC is stored in a built-in matrix data structure with a 2D
block cyclic distribution.  The Task Bench implementation divides this
matrix into $M$ by $N$ tiles where $N$ equals to the width of the task
graph and $M$ varies from 2 to the height of the task graph depending
on optimizations performed to alleviate write-after-read dependencies.
Tiles are distributed across nodes in pre-defined patterns,
and can be passed to tasks on other nodes, but
can only be written on the node where the data is originally
allocated. In our Task Bench implementation, each task writes to a
tile as output, and these tiles are passed as inputs to the
subsequent sets of tasks. As in the OmpSs and OpenMP implementations,
the task launch code uses a switch statement as the PaRSEC API assumes
tasks have a fixed number of dependencies.

@@

Realm~\cite{Realm14} is an explicitly-parallel task-based programming
model used internally by Legion~\cite{Legion12} and
Regent~\cite{Regent15}. Thus the Realm implementation of Task Bench can be
seen as a limit study of what can be achieved with Legion or Regent.

Tasks in Realm are executed in a deferred manner, with dependencies
between tasks explicitly defined by \emph{events} passed from one task
to another. Realm's data model supports collections that live in a
specific memory. Data must be explicitly copied, and the copies
synchronized with tasks via events.

@@

StarPU~\cite{StarPU11} is a task-based system that supports a \emph{sequential task flow} (STF)
programming model similar to PaRSEC's DTD. Our Task Bench implementation in
StarPU is very similar to PaRSEC, including the use of a switch
statement to handle dynamic numbers of task dependencies, and manual
elision of tasks on unrelated nodes to provide horizontal scalability.
A simplified version of the 2D block cyclic matrix from the Chameleon 
project~\cite{Chameleon} is used to describe the data produced by tasks in 
our Task Bench implementation.
